Cycle Route Uptake and Scenario Estimation (CRUSE): an open and nationally scalable approach for strategic cycle network planning tool development \\\large{A case study of Ireland}

### 2024-03-22 paper.qmd
Update paper.tex

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an open and nationally scalable approach for strategic cycle network planning tool development"
subtitle: A case study of Ireland
bibliography: bibliography.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  html: default
  # docx: default
# # # # Uncomment for formatted version
#   elsevier-pdf:
#     linestretch: 2
#     keep-tex: true
#     journal:
#       name: European Transport Research Review
#       formatting: preprint
#       model: 3p
#       cite-style: super
# # # #   # pdf: default

@@

::: {#fig-dublin}
<!-- ![](images/paste-1.png) -->
![](images/paste-8.png)
:::

<div>

<!-- ![](images/paste-3.png) -->
![](images/paste-9.png)

Illustration of the density of route networks generated by the CRUSE Tool for Dublin City and surroundings (top) and rural County Mayo on the West coast of Ireland (botton).
Source: CRUSE Tool, available at <https://cruse.bike/>.

</div>

>>>>

---
title: "Cycle Route Uptake and Scenario Estimation (CRUSE): an open and nationally scalable approach for strategic cycle network planning tool development"
subtitle: A case study of Ireland
bibliography: bibliography.bib
csl: apa.csl
editor:
  markdown:
    wrap: sentence
number-sections: true
latex-clean: false
pdf-engine: pdflatex
knitr:
  opts_chunk:
    fig.path: images/
format: 
  # html: default
  # docx: default
# # # # Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # # #   # pdf: default

@@

::: {#fig-dublin}
<!-- ![](images/paste-1.png) -->
![](images/paste-8.png)
<!-- ![](images/paste-3.png) -->
![](images/paste-9.png)

Illustration of the density of route networks generated by the CRUSE Tool for Dublin City and surroundings (top) and rural County Mayo on the West coast of Ireland (botton).
Source: CRUSE Tool, available at <https://cruse.bike/>.
:::

### 2024-03-22 paper.tex
Update paper.tex

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={CRUSE to Safe Cycling in Ireland},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{CRUSE to Safe Cycling in Ireland \\\large{An Open Source Tool to
Support Active Travel} }
\author[1]{Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Joey Talbot%
%
}

@@

\begin{abstract}
Legislation mandates improved road safety and decarbonisation of
predominantly road-based transport network in Ireland and other
countries. Increasing cycling transport, making it safer and more
attractive, is essential to this end. Cycling in Ireland represents only
3\% of total modal share as of the latest Census (2016), but accounts
for 20\% of serious injuries and 7\% of all fatalities. A planning tool
based on high resolution datasets on baseline and potential cycling
levels could greatly assist in achieving this goal.

This paper describes the design, features and potential use of the Cycle
Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland was
commissioned to support cycle network planning, safety interventions,
business cases and project evaluation. The resulting web application is
available at \url{https://cruse.bike/}, enabling planners, engineers,
and other stakeholders to make more evidence-based decisions. CRUSE goes
beyond previous work in several important ways. It models networks at
high levels of geographic resolution, simulates a wide range of trips
including recreational, social, shopping and personal trips, in addition
to trips to work and primary, secondary and tertiary education
locations. The tool provides key `denominator' information to enable
reporting of collision rates per billion km cycled.

Another key feature of CRUSE is its use of frequently-updated
OpenStreetMap data for route calculation and ``cycle friendliness''
estimates of all links on the network. CRUSE provides three network
types: the `Fastest' network highlights routes for directness;
`Quietest' prioritises cycle friendliness, and `Balanced' represents a
balance between speed and comfort. Trips are generated based on origin
and destination data from the 2016 Census in combination with modeled
demand data to estimate current and potential future cycling levels at
area, route, and network levels for each county in Ireland.

For investment in cycling to deliver on the huge potential benefits,
cycling must become safer. With growth in the ebike market, CRUSE will
help inform inter-urban and rural networks to support the transfer of
trips to sustainable modes for longer journeys. Provision of open access
data in a publicly available web tool will address the challenges faced
by counties across Ireland in meeting their sustainability commitments.
The experience of developing a deploying CRUSE are relevant to countries
with sustainable transport policies worldwide.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, sharp corners, frame hidden, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

@@

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

In Ireland, the Road Safety Authority (RSA) has set the target of
halving the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key active travel routes to be
identified and improved. A proactive approach to the development of
cycling policy and infrastructure is taking place in Ireland. The main
frameworks underpinning these efforts are the Climate Action Plan
\citep{climate2022a}, the National Development Plan, and the National
Roads 2040 strategy \citep{national2023}. These policy frameworks are
backed-up by public support for active travel interventions: while only
3\% of trips in Ireland are made by cycling according to the latest
Census data from 2016 ``83\% of respondents supported enhanced walking
and cycling facilities'' \citep{sustaina2019}.

@@

{\centering

\includegraphics{images/paste-1.png}

\includegraphics{images/paste-3.png}

@@

\caption{\label{fig-dublin}Illustration of the density of route networks
generated by the CRUSE Tool for Dublin City and surroundings (top) and
rural County Mayo on the West coast of Ireland. Source: CRUSE Tool,
available at \url{https://cruse.bike/}.}

\end{figure}

@@

\hypertarget{sec-methods}{%
\section{Methods and data}\label{sec-methods}}

@@

\hypertarget{sec-disaggregation}{%
\subsection{Disaggregation of origin-destination
data}\label{sec-disaggregation}}

@@

\hypertarget{sec-trip-purposes}{%
\subsection{Additional trip purposes}\label{sec-trip-purposes}}

@@

{\centering

@@

\end{figure}

\hypertarget{routing}{%
\subsection{Routing}\label{routing}}

@@

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}

\end{minipage}%
\newline
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}

\end{minipage}%

@@

\end{figure}

\hypertarget{sec-scenarios}{%
\subsection{Scenarios of cycling potential}\label{sec-scenarios}}

@@

\end{figure}

\hypertarget{baseline}{%
\subsubsection{Baseline}\label{baseline}}

@@

\hypertarget{near-market}{%
\subsubsection{Near Market}\label{near-market}}

@@

\hypertarget{climate-action-plan}{%
\subsubsection{Climate Action Plan}\label{climate-action-plan}}

@@

\hypertarget{go-dutch}{%
\subsubsection{Go Dutch}\label{go-dutch}}

@@

\hypertarget{ebike}{%
\subsubsection{Ebike}\label{ebike}}

@@

\hypertarget{sec-ui}{%
\subsection{User interface}\label{sec-ui}}

@@

\end{figure}

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

@@

\end{figure}

@@

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

@@

\hypertarget{sec-conclusions}{%
\section{Conclusions}\label{sec-conclusions}}

@@

\hypertarget{list-of-abbreviations}{%
\section{List of abbreviations}\label{list-of-abbreviations}}

@@

\hypertarget{declarations}{%
\section{Declarations}\label{declarations}}

\hypertarget{availability-of-data-and-material}{%
\subsection*{Availability of data and
material}\label{availability-of-data-and-material}}
\addcontentsline{toc}{subsection}{Availability of data and material}

@@

\hypertarget{funding}{%
\subsection*{Funding}\label{funding}}
\addcontentsline{toc}{subsection}{Funding}

@@

\hypertarget{acknowledgements}{%
\subsection*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{subsection}{Acknowledgements}

>>>>

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an open and nationally scalable approach for strategic cycle network planning tool development},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Tool; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{Cycle Route Uptake and Scenario Estimation (CRUSE): an open and
nationally scalable approach for strategic cycle network planning tool
development \\\large{A case study of Ireland} }
\author[1]{Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Joey Talbot%
%
}

@@

\begin{abstract}
This paper describes an approach to the development of strategic cycle
network planning tools. Based on our experience developing and deploying
the Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland,
we outline the methods used to generate the evidence presented in the
tool, including disaggregation of origin-destination data, additional
trip purposes, routing, scenarios of cycling potential, and the user
interface. The CRUSE Tool provides estimates of cycling currently and
under four scenarios of cycling potential: Near Market, Climate Action
Plan, Go Dutch, and Ebike. The tool was commissioned by Transport
Infrastructure Ireland and is publicly available at
\url{https://cruse.bike/}, enabling planners, engineers, and other
stakeholders to make more evidence-based decisions. CRUSE goes beyond
previous work by: modelling networks at high levels of geographic
resolution; simulating multiple trip purposes (social, shopping and
personal utility trips plus recreational and cycle touring trips), in
addition to using official data on trips to work and education
locations; provision of estimates of `quietness' (a proxy for
cyclability) at the route segment level; and its provision of three
network types: `Fastest', `Balanced' and `Quietest' to help plan both
arterial and residential cycle networks. Workshops and unstructure
interview is stakeholders show that the tool is already being used in
practice, to inform urban, inter-urban and rural network development.
The approach and methods for strategic cycle network development tools
outlined in this paper can be applied and built-on in other places to
support the generation of evidence needed to support ambitious cycling
policies and interventions internationally.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}

@@

\section{Introduction}\label{introduction}

@@

In Ireland, the Road Safety Authority (RSA) has set the target of
halving the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key active travel routes to be
identified and improved. Cycling in Ireland represents only 3\% of total
modal share as of the latest Census (2016), but accounts for 20\% of
serious injuries and 7\% of all fatalities. A proactive approach to the
development of cycling policy and infrastructure is taking place in
Ireland. The main frameworks underpinning these efforts are the Climate
Action Plan \citep{climate2022a}, the National Development Plan, and the
National Roads 2040 strategy \citep{national2023}. These policy
frameworks are backed-up by public support for active travel
interventions: while only 3\% of trips in Ireland are made by cycling
according to the latest Census data from 2016 ``83\% of respondents
supported enhanced walking and cycling facilities''
\citep{sustaina2019}.

@@

\centering{

\includegraphics{images/paste-8.png}
\includegraphics{images/paste-9.png}

@@

\caption{\label{fig-dublin}Illustration of the density of route networks
generated by the CRUSE Tool for Dublin City and surroundings (top) and
rural County Mayo on the West coast of Ireland (botton). Source: CRUSE
Tool, available at \url{https://cruse.bike/}.}

\end{figure}%

@@

\section{Methods and data}\label{sec-methods}

@@

\subsection{Disaggregation of origin-destination
data}\label{sec-disaggregation}

@@

\subsection{Additional trip purposes}\label{sec-trip-purposes}

@@

\centering{

@@

\end{figure}%

\subsection{Routing}\label{routing}

@@

\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}\end{minipage}%
\newline
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}\end{minipage}%
%
\begin{minipage}{0.50\linewidth}
\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}\end{minipage}%

@@

\end{figure}%

\subsection{Scenarios of cycling potential}\label{sec-scenarios}

@@

\end{figure}%

\subsubsection{Baseline}\label{baseline}

@@

\subsubsection{Near Market}\label{near-market}

@@

\subsubsection{Climate Action Plan}\label{climate-action-plan}

@@

\subsubsection{Go Dutch}\label{go-dutch}

@@

\subsubsection{Ebike}\label{ebike}

@@

\subsection{User interface}\label{sec-ui}

@@

\end{figure}%

\section{Results}\label{sec-results}

@@

\end{figure}%

@@

\section{Discussion}\label{sec-discussion}

@@

\section{Conclusions}\label{sec-conclusions}

@@

\section{List of abbreviations}\label{list-of-abbreviations}

@@

\section{Declarations}\label{declarations}

\subsection*{Availability of data and
material}\label{availability-of-data-and-material}
\addcontentsline{toc}{subsection}{Availability of data and material}

@@

\subsection*{Funding}\label{funding}
\addcontentsline{toc}{subsection}{Funding}

@@

\subsection*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{subsection}{Acknowledgements}
